Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-13 aistats.tex
fixed most of the warnings

\documentclass[twoside]{article}
\usepackage{aistats/aistats2022}
% If your paper is accepted, change the options for the package
% aistats2022 as follows:
%\usepackage[accepted]{aistats2022}

@@

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{{figs/}}
%\usepackage{subfigure}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}       % hyperlinks
\usepackage[dvipsnames]{xcolor}

@@

\hypersetup{ % SLJ: my standard paper setup...
	pdftitle={MAP convergence},
	pdfkeywords={},
	pdfborder=0 0 0,
	pdfpagemode=UseNone,
	colorlinks=true,
	linkcolor=blue, %mydarkblue,
	citecolor=blue, %mydarkblue,
	filecolor=blue, %mydarkblue,
	urlcolor=blue, %mydarkblue,
	pdfview=FitH,
	pdfauthor={Anonymous},
}

@@

\hypersetup{draft}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{{figs/gaussian_variance.pdf}}
	\caption{
	\textbf{Gaussian variance $\cN(0,\sigma^2)$ example. Left:} training curves and analytic upper bound.
	\textbf{Center:} bias-mixed-variance decomposition, using the arithmetic mean.
	\textbf{Right:} bias-variance decomposition, using the harmonic mean.
	}
	\label{fig:variance_decomposition}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{{figs/gaussians/new_linear_n0=1}.pdf}
	\caption{
	\textbf{Full gaussian} $\cN(m, \sigma^2)$ with $\meanp^*=(0, 1), \meanp_0 = (1,2)$ and $n_0=1$. \textbf{Left:} training curves (5,95) percentiles, average and asymptote.
	\textbf{Right:} bias-variance decomposition.
	}
	\label{fig:gaussian_decomposition}
\end{figure}

@@

\paragraph{Online Learning}
Other works focus instead on the online estimation of $\theta$ and minimize the \textit{regret}
\[
R_T=\sum_{t=1}^T f_t(\hat{\theta}_t)-\min_{\theta\in\Theta}\sum_{t=1}^Tf_t(\theta),
\]
where $f_t(\theta)\triangleq-\log p(x_t|\theta)$, and the estimator $\hat{\theta}_t$ is computed using only the $t-1$ samples $\{x_i\}$. Usually, the \textit{follow-the-leader} strategy works best, i.e., moving the current estimate towards $x_i$ with a certain step size schedule. Unfortunately however, their rate are worse than the offline MAP or MLE algorithms. For instance, in the specific case where we estimate the mean parameter of a Bernoulli or Gaussian distribution, \citet{azoury2001relative} show that $R_T\leq log(T)$. From this bound, we can deduce the rate  to the real mean parameter, being $f(\hat{\theta}_T)-f^*\leq \frac{log(T)}{T}$. Also, \citet{dasgupta2007online} proves the asymptotic rate $\lim\sup_{T>1}\frac{R_T}{T}\leq 0$ when we estimate the mean and variance of a Gaussian, under the (strong) assumption that the data are finite, i.e., $\|x_t\|<C$. This means the estimator converges asymptotically, but with an unknown rate of convergence, therefore highlighting the difficulty of that setting.

@@

\newpage
\bibliographystyle{apalike}
\bibliography{references.bib}

>>>>

\documentclass[twoside]{article}

\usepackage{silence}
% Disable all warnings issued by latex starting with "You have..."
% This is a hack, there is no clean solution for that unless we install locally the package
\WarningFilter{latex}{You have requested package}

\pdfsuppresswarningpagegroup=1 %removes the warning
% This is a hack, the other way to remove is to convert all pdf into postscript or eps.

\usepackage{aistats/aistats2022}
% If your paper is accepted, change the options for the package
% aistats2022 as follows:
%\usepackage[accepted]{aistats2022}

@@

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{{figs/}}
%\usepackage{subfigure}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}       % hyperlinks
\usepackage[dvipsnames]{xcolor}

@@

\hypersetup{ % SLJ: my standard paper setup...
	pdftitle={MAP convergence},
	pdfkeywords={},
	pdfborder=0 0 0,
	pdfpagemode=UseNone,
	colorlinks=true,
	linkcolor=blue, %mydarkblue,
	citecolor=blue, %mydarkblue,
	filecolor=blue, %mydarkblue,
	urlcolor=blue, %mydarkblue,
	pdfview=FitH,
	pdfauthor={Anonymous},
	% draft,
	final,
}

@@

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{figs/gaussian_variance.pdf}
	\caption{
	\textbf{Gaussian variance $\cN(0,\sigma^2)$ example. Left:} training curves and analytic upper bound.
	\textbf{Center:} bias-mixed-variance decomposition, using the arithmetic mean.
	\textbf{Right:} bias-variance decomposition, using the harmonic mean.
	}
	\label{fig:variance_decomposition}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{figs/gaussians/new_linear_n0=1.pdf}
	\caption{
	\textbf{Full gaussian} $\cN(m, \sigma^2)$ with $\meanp^*=(0, 1), \meanp_0 = (1,2)$ and $n_0=1$. \textbf{Left:} training curves (5,95) percentiles, average and asymptote.
	\textbf{Right:} bias-variance decomposition.
	}
	\label{fig:gaussian_decomposition}
\end{figure}

@@

\paragraph{Online Learning}
Other works focus instead on the online estimation of $\theta$ and minimize the \textit{regret}
\[
R_T=\sum_{t=1}^T f_t(\hat{\theta}_t)-\min_{\theta\in\Theta}\sum_{t=1}^Tf_t(\theta),
\]
where $f_t(\theta)\triangleq-\log p(x_t|\theta)$, and the estimator $\hat{\theta}_t$ is computed using only the $t-1$ samples $\{x_i\}$. Usually, the \textit{follow-the-leader} strategy works best, i.e., moving the current estimate towards $x_i$ with a certain step size schedule. Unfortunately however, their rate are worse than the offline MAP or MLE algorithms. For instance, in the specific case where we estimate the mean parameter of a Bernoulli or Gaussian distribution, \citet{azoury2001relative} show that $R_T\leq log(T)$. From this bound, we can deduce the rate  to the real mean parameter, being $f(\hat{\theta}_T)-f^*\leq \frac{log(T)}{T}$. Also, \citet{dasgupta2007online} prove the asymptotic rate $\lim\sup_{T>1}\frac{R_T}{T}\leq 0$ when we estimate the mean and variance of a Gaussian, under the (strong) assumption that the data are finite, i.e., $\|x_t\|<C$. This means the estimator converges asymptotically, but with an unknown rate of convergence, therefore highlighting the difficulty of that setting.

@@

\newpage
\bibliographystyle{apalike}
\bibliography{references} % Removed the .bib as this mess up with the compiler
